### 2019-11-08 paper/conceptual_framework.tex
Add contraception and spacing discussion

The decisions on sexual activity and contraceptive effort are constrained by biological 
factors, such as fecundity, age, and health of the mother, and the knowledge of and access
to contraceptive technology.
There is evidence from ``natural fertility'' populations that increasing number of 
children is associated with shorter spacing and that birth intervals increase 
with parity \citep{Henry1961,Leridon1977}.

@@

TK contraception

There is no clear relationship between contraception use and birth spacing in the 
literature, and better-educated women have shorter spacing than less educated women in 
some instances but not in others 
\citep{Tulasidhar1993,Whitworth2002,Bhalotra2008,Yeakey2009,Kim2010,Soest2018}.

Even the effects of access to contraceptive and declining fertility on spacing are unclear.
On the one hand, access to contraceptives allows women to avoid too short spacing between 
birth.
On the other hand, increased reliability of access and effectiveness of contraceptives can 
lead to shorter intervals between births if women used to have longer spacing to avoid
having too many children by accident \citep{Keyfitz1971,Heckman1976}.
%
\footnote{
In Matlab, the effect of son preference on birth spacing is stronger in areas with better 
access to family planning \citep{Rahman1993}.
Other evidence, however, points to families ability to time births, even in the absence
of modern contraceptives \citep{Jayachandran2011,Alam2018}.
}

@@

>>>>

The decisions on sexual activity and contraceptive effort are moderated by fecundity,
determined by biological factors such as age and health of the mother, and the knowledge of and 
access to contraceptive technology.
There is evidence from ``natural fertility'' populations that increasing number of 
children is associated with shorter spacing and that birth intervals increase 
with parity \citep{Henry1961,Leridon1977}.

@@

Empirically, there is no clear relationship between contraception use and birth 
spacing or between education level and birth spacing
\citep{Tulasidhar1993,Whitworth2002,Bhalotra2008,Yeakey2009,Kim2010,Soest2018}.
The lack of evidence is consistent with the ambiguous theoretical effect of introducing 
modern contraceptives on birth spacing.
On one hand, when contraceptive technologies are less effective, 
parents may try to avoid overshooting their desired number of children by spacing their 
children further apart.
Increased reliability of access to and effectiveness of contraceptives can then lead to 
shorter intervals between births because parents are more certain of effectively
stopping childbearing when they reach their preferred number of children and having
the children closer together provide higher utility because parents get to enjoy the
children for longer as discussed above \citep{Keyfitz1971,Heckman1976}.
On the other hand, when contraceptives are less effective we would expect, through 
random variation, some birth intervals to be shorter than ideal for the reasons 
discussed below.
With increased reliability of access to and effectiveness of contraceptives, parents 
would better be able to avoid too short spacing between birth and average spacing
should go up.

To further complicate the relationship between contraception access and spacing,
the ability to successfully use ``low efficacy'' contraceptive methods, such as 
the rhythm method, is increasing in education, while there is no difference across education 
leves when it comes to modern contraceptives such as pills or injectables 
\citep{Rosenzweig1989}.
Furthermore, even in the absence of modern contraceptives, families appear able to time or 
at least postpone births when faced with strong enough incentives 
\citep{Jayachandran2011,Alam2018}.
[Does this need a tie-in to the paragraph above?]

@@

In Matlab, the effect of son preference on birth spacing is stronger in areas with better 
access to family planning \citep{Rahman1993}.
